Neutral models of \textit{de novo} gene emergence show that gene evolution has a preferred trajectory

### 2022-11-10 Manuscripts/deNovoRates.tex
modified:   Manuscripts/deNovoRates.tex
modified:   Manuscripts/deNovoRates.tex 	modified:   Manuscripts/refs.bib

The second major requirement for protein expression is translation of the transcribed mRNA. The most fundamental requirement for translation is the presence of an open reading frame (ORF). Usually, an mRNA needs additional features to initiate protein synthesis. These features include ribosome binding sites in prokaryotes \citep{RBS}, and Kozak consensus sequences \citep{kozak} in eukaryotes.

@@

The consensus sequence of TATA-box is \texttt{TATA [T/A] A [T/A] [G/A]} \citep{tata2} such that, any of the nucleotides enclosed within the square brackets, can exist in that specific position in the sequence. Another study used X-ray crystallography to identify mutations that are tolerated in TATA-box consensus sequence \citep{tata1}. Thus there are eight consensus sequences, and at seven positions, a non-consensus nucleotide is tolerated (\autoref{tatatab}). This makes a TATA-box feature set that contains 136 sequences. We defined the corresponding non-feature set by generating all 8-mer DNA sequences that do not belong to the TATA-box feature set.

\begin{table}[H]
\centering
\begin{tabular}{@{}ccc@{}}
\toprule
Position  & Accepted nucleotides & \begin{tabular}[c]{@{}c@{}}Number of Variants\\ (consensus + suboptimal)\end{tabular} \\ \midrule
1         & T$\gg$C$>$A,G        & $1+3$                                                                                 \\ \midrule
2         & A$\gg$T              & $1+1$                                                                                 \\ \midrule
3         & T$\gg$A,C            & $1+2$                                                                                 \\ \midrule
4         & A$\gg$T              & $1+1$                                                                                 \\ \midrule
5         & T/A                  & $2+0$                                                                                 \\ \midrule
6         & A$\gg$G$>$C/T        & $1+3$                                                                                 \\ \midrule
7         & A/T$>$G$>$C          & $2+2$                                                                                 \\ \midrule
8         & G/A$>$C/T            & $2+2$                                                                                                                                                                 \\ \bottomrule
\end{tabular}
\caption{Nucleotide residues that are tolerated at different positions in a TATA-box \citep[based on][]{tata1}}
\label{tatatab}
\end{table}

>>>>

The second major requirement for protein expression is translation of the transcribed mRNA. The most fundamental requirement for translation is the presence of an open reading frame (ORF). Usually, an mRNA needs additional features to initiate protein synthesis. These features include ribosome binding sites in prokaryotes \citep{RBS}, and Kozak consensus sequences in eukaryotes \citep{kozak,kozakDroso,kozakHuman}.

@@

The consensus sequence of TATA-box is \texttt{TATA [T/A] A [T/A] [G/A]} \citep{tata2} such that, any of the nucleotides enclosed within the square brackets, can exist in that specific position in the sequence. Another study used X-ray crystallography to identify mutations that are tolerated in TATA-box consensus sequence \citep{tata1}. Thus there are eight consensus sequences, and at seven positions, a non-consensus nucleotide is tolerated (\autoref{tatatab}). This makes a TATA-box feature set that contains 104 sequences. We defined the corresponding non-feature set by generating all 8-mer DNA sequences that do not belong to the TATA-box feature set.

\begin{table}[H]
\centering
\begin{tabular}{c c}
\toprule
Position  & Accepted nucleotides \\ \midrule
1         & T$\gg$C$>$A,G        \\ \midrule
2         & A$\gg$T               \\ \midrule
3         & T$\gg$A,C             \\ \midrule
4         & A$\gg$T               \\ \midrule
5         & T/A                   \\ \midrule
6         & A$\gg$G$>$C/T         \\ \midrule
7         & A/T$>$G$>$C           \\ \midrule
8         & G/A$>$C/T             \\ \bottomrule
\end{tabular}
\caption{Nucleotide residues that are tolerated at different positions in a TATA-box \citep[based on][]{tata1}}
\label{tatatab}
\end{table}
